---
output: html_document
editor_options: 
  chunk_output_type: console
---
# IE - Master in Business Analytics and Big Data
Recommendation Systems

Building a Non Personalized Recommendation Engine

Author: Cris Pineda

========================================================

# Initialization

Import libraries

```{r}
library(dplyr)
library(ggplot2)
library(recommenderlab)
library(data.table)
library(reshape2)
```

Load the critics dataset.

```{r}
critics <- read.csv("/Users/cris/Desktop/Courses/Second Semester/Reccomendation Engines/IE-Lab1-Package/Data/critics.csv")
View(critics)
```

# 1) Top 5 Movies, by rating mean
Calculate mean rating for each movie, ordered with the highest rating listed first, and plot the top 5.

**Expected result**:
3.600000, 'The Shawshank Redemption'
3.266667, 'Star Wars IV - A New Hope'
3.222222, 'Blade Runner'
3.166667, 'Groundhog Day'
3.062500, 'The Silence of the Lambs'

```{r fig.width=7, fig.height=6}
# use colMeans to get the mean rating of each movie (remove first row) and sort by highest ratings and limit to 5
top5_ratings_mean <- head(sort(colMeans(critics[,-1], na.rm = T), decreasing = T), n=5)
```

# 2) Top 5 Movies, by rating distribution
Calculate the percentage of ratings for each movie that are 4 stars or higher. Order with the highest percentage first, and plot the top 5 movies with more high ratings distribution.

**Expected result**:
0.7000000, 'The Shawshank Redemption'
0.5333333, 'Star Wars IV - A New Hope'
0.5000000, 'Gladiator'
0.4444444, 'Blade Runner'
0.4375000, 'The Silence of the Lambs'

```{r fig.width=7, fig.height=6}
# create an empty list for the movies
movie_list <- list()


# Star.Wars.IV...A.New.Hope
Star.Wars.IV <- critics %>%
  select("Star.Wars.IV...A.New.Hope") # select only this row
SWIV <- (sum(Star.Wars.IV == 4, na.rm = T)+sum(Star.Wars.IV == 5, na.rm = T))/sum(Star.Wars.IV>0, na.rm = T) # create new variable to get percentage
movie_list <- append(movie_list, c("Star.Wars.IV...A.New.Hope", SWIV)) # append to the empty list

## repeat for each movie :(

# Star.Wars.VI...Return.of.the.Jedi
Star.Wars.VI <- critics %>%
  select("Star.Wars.VI...Return.of.the.Jedi")
SWVI <- (sum(Star.Wars.VI == 4, na.rm = T)+sum(Star.Wars.VI == 5, na.rm = T))/sum(Star.Wars.VI>0, na.rm = T)
movie_list <- append(movie_list, c("Star.Wars.VI...Return.of.the.Jedi", SWVI))

# Forrest.Gump
Forrest.Gump <- critics %>%
  select("Forrest.Gump")
FG <- (sum(Forrest.Gump == 4, na.rm = T)+sum(Forrest.Gump == 5, na.rm = T))/sum(Forrest.Gump>0, na.rm = T)
movie_list <- append(movie_list, c("Forrest.Gump", FG))

# The.Shawshank.Redemption
The.Shawshank.Redemption <- critics %>%
  select("The.Shawshank.Redemption")
TSR <- (sum(The.Shawshank.Redemption == 4, na.rm = T)+sum(The.Shawshank.Redemption == 5, na.rm = T))/sum(The.Shawshank.Redemption>0, na.rm = T)
movie_list <- append(movie_list, c("The.Shawshank.Redemption", TSR))

# The.Silence.of.the.Lambs
The.Silence.of.the.Lambs <- critics %>%
  select("The.Silence.of.the.Lambs")
TSOTL <- (sum(The.Silence.of.the.Lambs == 4, na.rm = T)+sum(The.Silence.of.the.Lambs == 5, na.rm = T))/sum(The.Silence.of.the.Lambs>0, na.rm = T)
movie_list <- append(movie_list, c("The.Silence.of.the.Lambs", TSOTL))

# Gladiator
Gladiator <- critics %>%
  select("Gladiator")
GLAD <- (sum(Gladiator == 4, na.rm = T)+sum(Gladiator == 5, na.rm = T))/sum(Gladiator>0, na.rm = T)
movie_list <- append(movie_list, c("Gladiator", GLAD))

# Toy.Story
Toy.Story <- critics %>%
  select("Toy.Story")
TS <- (sum(Toy.Story == 4, na.rm = T)+sum(Toy.Story == 5, na.rm = T))/sum(Toy.Story>0, na.rm = T)
movie_list <- append(movie_list, c("Toy.Story", TS))

# Saving.Private.Ryan
Saving.Private.Ryan <- critics %>%
  select("Saving.Private.Ryan")
SPR <- (sum(Saving.Private.Ryan == 4, na.rm = T)+sum(Saving.Private.Ryan == 5, na.rm = T))/sum(Saving.Private.Ryan>0, na.rm = T)
movie_list <- append(movie_list, c("Saving.Private.Ryan", SPR))

# Pulp.Fiction
Pulp.Fiction <- critics %>%
  select("Pulp.Fiction")
PF <- (sum(Pulp.Fiction == 4, na.rm = T)+sum(Pulp.Fiction == 5, na.rm = T))/sum(Pulp.Fiction>0, na.rm = T)
movie_list <- append(movie_list, c("Pulp.Fiction", PF))

# Stand.by.Me
Stand.by.Me <- critics %>%
  select("Stand.by.Me")
SBM <- (sum(Stand.by.Me == 4, na.rm = T)+sum(Stand.by.Me == 5, na.rm = T))/sum(Stand.by.Me>0, na.rm = T)
movie_list <- append(movie_list, c("Stand.by.Me", SBM))

# Shakespeare.in.Love
Shakespeare.in.Love <- critics %>%
  select("Shakespeare.in.Love")
SIL <- (sum(Shakespeare.in.Love == 4, na.rm = T)+sum(Shakespeare.in.Love == 5, na.rm = T))/sum(Shakespeare.in.Love>0, na.rm = T)
movie_list <- append(movie_list, c("Shakespeare.in.Love", SIL))

# Total.Recall
Total.Recall <- critics %>%
  select("Total.Recall")
TR <- (sum(Total.Recall == 4, na.rm = T)+sum(Total.Recall == 5, na.rm = T))/sum(Total.Recall>0, na.rm = T)
movie_list <- append(movie_list, c("Total.Recall", TR))

# Independence.Day
Independence.Day <- critics %>%
  select("Independence.Day")
ID <- (sum(Independence.Day == 4, na.rm = T)+sum(Independence.Day == 5, na.rm = T))/sum(Independence.Day>0, na.rm = T)
movie_list <- append(movie_list, c("Independence.Day", ID))

# Blade.Runner
Blade.Runner <- critics %>%
  select("Blade.Runner")
BR <- (sum(Blade.Runner == 4, na.rm = T)+sum(Blade.Runner == 5, na.rm = T))/sum(Blade.Runner>0, na.rm = T)
movie_list <- append(movie_list, c("Blade.Runner", BR))

# Groundhog.Day
Groundhog.Day <- critics %>%
  select("Groundhog.Day")
GD <- (sum(Groundhog.Day == 4, na.rm = T)+sum(Groundhog.Day == 5, na.rm = T))/sum(Groundhog.Day>0, na.rm = T)
movie_list <- append(movie_list, c("Groundhog.Day", GD))

# The.Matrix
The.Matrix <- critics %>%
  select("The.Matrix")
TM <- (sum(The.Matrix == 4, na.rm = T)+sum(The.Matrix == 5, na.rm = T))/sum(The.Matrix>0, na.rm = T)
movie_list <- append(movie_list, c("The.Matrix", TM))

# Schindler.s.List
Schindler.s.List <- critics %>%
  select("Schindler.s.List")
SL <- (sum(Schindler.s.List == 4, na.rm = T)+sum(Schindler.s.List == 5, na.rm = T))/sum(Schindler.s.List>0, na.rm = T)
movie_list <- append(movie_list, c("Schindler.s.List", SL))


# The.Sixth.Sense
The.Sixth.Sense <- critics %>%
  select("The.Sixth.Sense")
TSS <- (sum(The.Sixth.Sense == 4, na.rm = T)+sum(The.Sixth.Sense == 5, na.rm = T))/sum(The.Sixth.Sense>0, na.rm = T)
movie_list <- append(movie_list, c("The.Sixth.Sense", TSS))

# Raiders.of.the.Lost.Ark
Raiders.of.the.Lost.Ark <- critics %>%
  select("Raiders.of.the.Lost.Ark")
ROTLA <- (sum(Raiders.of.the.Lost.Ark == 4, na.rm = T)+sum(Raiders.of.the.Lost.Ark == 5, na.rm = T))/sum(Raiders.of.the.Lost.Ark>0, na.rm = T)
movie_list <- append(movie_list, c("Raiders.of.the.Lost.Ark", ROTLA))

# Babe
Babe <- critics %>%
  select("Babe")
B <- (sum(Babe == 4, na.rm = T)+sum(Babe == 5, na.rm = T))/sum(Babe>0, na.rm = T)
movie_list <- append(movie_list, c("Babe", B))

# turn list into data frame and name columns
movie_df <- data.frame(matrix(unlist(movie_list), ncol=2, byrow=T))
colnames(movie_df) <- c("Movie","Percentage")

# get the top 5
movie_df %>%
  arrange(desc(Percentage)) %>%
  head(5)
```

# 3) Top 5 Movies, by quantity of ratings
Count the number of ratings for each movie, order the one with most number of ratings first, submit the top 5.

**Expected result**:
17, 'Toy Story'
16, 'The Silence of the Lambs'
15, 'Star Wars IV - A New Hope'
14, 'Star Wars VI - Return of the Jedi'
13, 'Independence Day'

```{r fig.width=7, fig.height=6}
# get the total row by nrow and by using sapply, minus the na rows so you get how many ratings each movie has. Then sort by the highest and limit it to 5.
top5_ratings_count <- head(sort(nrow(critics)-colSums(sapply(critics[,-1],is.na)), decreasing =TRUE), n=5)
```

# 4) People who watched Star Wars IV also watched ...
Calculate movies that most often occur with other movie. For each movie, calculate the percentage of the other movie raters who also rated that movie. Order with the highest percentage first, and submit the top 5. Select "Star.Wars.IV...A.New.Hope" movie by defaut. Use the (x+y)/x method.

Top Occurences:
0.9333333, 'Toy Story'
0.8666667, 'Star Wars: Episode VI - Return of the Jedi'
0.8000000, 'The Silence of the Lambs'
0.7333333, 'Independence Day'
0.6666667, 'Total Recall'

```{r fig.width=7, fig.height=6}
# remove all the critics who did not rate Star War IV
StarWars_fans <- filter(critics, Star.Wars.IV...A.New.Hope>0)
# remove the first two columns since they are not needed
StarWars_fans <- StarWars_fans[,-1:-2]

# get the percentage of occurences (might be an easier way but idk) and then sort by highest and limit to 4
head(sort((nrow(StarWars_fans)-colSums(is.na(StarWars_fans))) / nrow(StarWars_fans), decreasing = T),n=5)
```


# 5) People who liked Babe also liked ...
Calculate the movies better rated of people who liked a movie. Select the people who liked the movie "Babe" (4 or 5 stars) and provide the top 5 movies they liked most.

**Expected result**:
5.000, 'Pulp.Fiction'
5.000, 'Groundhog.Day'
4.500, 'The.Shawshank.Redemption'
4.333, 'Toy.Story'
4.000, 'Blade.Runner'

```{r fig.width=7, fig.height=6}
# filter dataset by users who rated "Babe" 4 or 5 stars
Babe_lovers <- filter(critics, Babe == 4 | Babe == 5)

# use colMeans to get the mean of each movie from the filtered data set and set the results to the top 5
Babe_lovers <- head(sort(colMeans(Babe_lovers[,-1], na.rm = T), decreasing = T), n=5)

Babe_lovers
```

# 6) MovieLens
Explore a real data set and provide non-personalized ratings. You can use the MovieLens database. You can find MovieLens' dataset here: http://files.grouplens.org/datasets/movielens/ml-10m-README.html

```{r}
# load both data sets
real_movie <- read.csv("/Users/cris/Desktop/Courses/Second Semester/Reccomendation Engines/IE-Lab1-Package/Data/ml-latest-small/movies.csv")
real_rating <- read.csv("/Users/cris/Desktop/Courses/Second Semester/Reccomendation Engines/IE-Lab1-Package/Data/ml-latest-small/ratings.csv")

# merge two data frames by ID
MovieLens <- merge(real_movie,real_rating,by="movieId")
MovieLens$rating <- as.numeric(MovieLens$rating)

# check for top 10 rated (quantity) movies
mostRated <- MovieLens %>%
  group_by(movieId) %>%
  summarise(qt = n()) %>%
  arrange(desc(qt)) %>%
  head(10)

# filter MovieLens
MovieLens <- MovieLens %>%
  select(movieId, title, userId, rating) %>%
  filter(movieId %in% c(356, 296, 318, 593, 260, 480, 2571, 1, 527, 589))

# now get rid of movieId
MovieLens <- MovieLens[,-1]

# use reshape to convert it to similar format of the exercise
MovieLens <- dcast(MovieLens, title ~ userId)

# we dont need all of them for the sample
MovieLens <- MovieLens[ ,1:21]

# transpose it
MovieLens <- t(MovieLens)

# fix the column names again
colnames(MovieLens) <- MovieLens[1,]
MovieLens <- MovieLens[-1,]

MovieLens1 <- as.numeric(as.character(MovieLens))

# having multiple errors trying to convert MovieLens to numeric so I will export it, change it on DataIku and bring it back
write.csv(MovieLens, file = "MovieLens.csv")

MovieLens <- read.csv("/Users/cris/MovieLens.csv", header = T)

# fix the row names again
rownames(MovieLens) <- MovieLens[,1]
MovieLens <- MovieLens[,-1]
```

Now we give some recommendations

# Top 5 Ratings by Mean
```{r}
ML5 <- head(sort(colMeans(MovieLens, na.rm = T), decreasing = T), n=3)

ML5
```

The top 3 reccomendations are The Shawshank Redemption, The Matrix, and Terminator 2: Judgmement Day.


For users who liked Schindler's List, what movie they would like...
```{r fig.width=7, fig.height=6}
# filter dataset by users who rated "Babe" 4 or 5 stars
Schindlers_Lovers <- filter(MovieLens, Schindler.s.List..1993. == 4 | Schindler.s.List..1993. == 5)

Schindlers_Lovers <- head(sort(colMeans(Schindlers_Lovers, na.rm = T), decreasing = T), n=1)

Schindlers_Lovers
```

Users who liked Schindler's List would probably like The Matrix most from these movies.


Let us see for Jurassic Park
```{r fig.width=7, fig.height=6}
# filter dataset by users who rated "Babe" 4 or 5 stars
JP_lovers <- filter(MovieLens, Jurassic.Park..1993. == 4 | Jurassic.Park..1993. == 5)

JP_lovers <- head(sort(colMeans(JP_lovers, na.rm = T), decreasing = T), n=1)

JP_lovers
```

Users who like Jurassic Park would probably like Pulp Fiction most from these movies.


Okay let us see for users who liked Forrest Gump, which movie they would probably like the least out of these top 10.

```{r fig.width=7, fig.height=6}
# filter dataset by users who rated "Babe" 4 or 5 stars
haters <- filter(MovieLens, Forrest.Gump..1994. == 4 | Forrest.Gump..1994. == 5)

# use colMeans to get the mean of each movie from the filtered data set and set the results to the top 5
haters <- head(sort(colMeans(haters, na.rm = T), decreasing = F), n=1)

haters
```

Users who liked Forrest Gump would probably like The Silence of the Lambs the least out of these movies.