---
output: html_document
editor_options: 
  chunk_output_type: console
---
# IE - Master in Business Analytics and Big Data
Recommendation Systems

Building a Collaborative Filtering Engine

Author: FirstName LastName

========================================================

# Initialization

Import libraries

```{r}
library(dplyr)
library(ggplot2)
library(recommenderlab)
library(data.table)
library(reshape2)
library(Hmisc)
```

Load the critics dataset.

```{r}
critics <- read.csv("/Users/cris/Desktop/Courses/Second Semester/Reccomendation Engines/IE-Lab1-Package/Data/critics.csv", header = T)
View(critics)
```

# 1) Pearson correlation coefficient 
Calculate the Pearson correlation coefficient between Victoria and Nuria

**Expected correlation between Victoria and Nuria**: 0.3651484

```{r fig.width=7, fig.height=6}
# transpose the data frame
critics_t <- data.frame(t(critics))

# turn the first row into the header
critics_t <- critics_t[-1, ] # removing the first row.
colnames(critics_t) = c("John", "Maria", "Anton", "Roger", "Martina", "Ana", "Sergi", "Marc", "Jim", "Chris", "Bernard", "Nuria", "Nerea", "Carles", "Victoria", "Ivan", "Rachel", "Nadia", "Oriol", "Valery") # assign names to headers

# create the correlation matrix using the Hmisc library
critics_t_corr <- rcorr(as.matrix(critics_t, type="pearson"))
critics_t_corr = critics_t_corr$r

# filter the correlation between Victoria and Nuria
VicNur <- critics_t_corr["Victoria", 'Nuria']

cat("Expected correlation between Victoria and Nuria:", VicNur)
```
# 2) Compare two movie critics
Using the calculated Pearson correlation in the previous section. Compare and show the movie ratings for Victoria and Nuria on a scatter plot. How is this compared with the same ratings of Maria and Nerea on a scatter plot? What about Chris and Jim? What it means?

```{r}
# plot Victoria and Nuria
ggplot(data = critics_t) +
  geom_point(aes(x = critics_t[, "Victoria"], y = critics_t[, "Nuria"], color=I("blue")))
cat("Expected correlation between Victoria and Nuria:", critics_t_corr["Victoria", 'Nuria'])

# plot Maria and Nerea
ggplot(data = critics_t) +
  geom_point(aes(x = critics_t[, "Maria"], y = critics_t[, "Nerea"], color=I("blue")))
cat("Expected correlation between Maria and Nerea:", critics_t_corr["Maria", 'Nerea'])

# plot Chris and Jim
ggplot(data = critics_t) +
  geom_point(aes(x = critics_t[, "Chris"], y = critics_t[, "Jim"], color=I("blue")))
cat("Expected correlation between Chris and Jim:", critics_t_corr["Chris", 'Jim'])
```

Victoria and Nuria have a higher corrleation than Maria and Nerea, but Chris and Jim have a higher correlation than Victoria and Nuria with a value of 0.8255243.

# 3) Top Recommendations 
Get the top 5 movies recommended for Victoria. by using a weighted average of every other critic's rankings.

**Expected result**:
'The Matrix'
'Forrest Gump'     
'The Sixth Sense' 
'Shakespeare in Love'
'Blade Runner'


```{r}
# use correlations and convert to numeric
critics_tn <- data.frame(lapply(critics_t, as.numeric))
critics_tn$movie <- rownames(critics_t) # add row names that were deleted
rownames(critics_tn) <- critics_tn[,21] # turn that column to row names
critics_tn <- critics_tn[,-21] # get rid of it now

# get mean for each User
cmeans <- sapply(critics_tn, mean, na.rm = T)

# movies Victoria has not seen
noVic <- rownames(critics_tn[is.na(critics_tn$Victoria == "NA"), 15, drop=FALSE])
noVic <- critics_tn[is.na(critics_tn$Victoria == "NA"), ]
#sumcols <- colSums(corrdf, na.rm = T)

# add the means to the noVic dataframe
newcritics <- cbind(critics_tn, cmeans)

# corrleation for Victoria
Victoria <- critics_t_corr[-15, "Victoria"]

# save correlations
cJohn = 0.27709796
cMaria = 0.49259218
cAnton = -0.18466410
cRoger = -0.30698876
cMartina = 0.04771653
cAna = 0.59761430
cSergi = 0.03636965
cMarc = 0.07934228
cJim = -0.44451445
cChris = -0.34426519
cBernard = -0.18649966
cNuria = 0.36514837
cNerea = 0.20573780
cCarles = 0.42732467
cIvan = -0.73277672
cRachel = 0.94491118
cNadia = -0.58737848
cOriol = 0.57897945
cValery = -0.08811342

# sort correlations by highest
sort(Victoria, decreasing = T)

# create the correlation df
corrdf <- noVic %>%
  mutate(John = (John - cmeans["John"]) * cJohn,
         Maria = (Maria - cmeans["Maria"]) * cMaria,
         Anton = (Anton - cmeans["Anton"]) * cAnton,
         Roger = (Roger - cmeans["Roger"]) * cRoger,
         Martina = (Martina - cmeans["Martina"]) * cMartina,
         Ana = (Ana - cmeans["Ana"]) * cAna,
         Sergi = (Sergi - cmeans["Sergi"]) * cSergi,
         Marc = (Marc - cmeans["Marc"]) * cMarc,
         Jim = (Jim - cmeans["Jim"]) * cJim,
         Chris = (Chris - cmeans["Chris"]) * cChris,
         Bernard = (Bernard - cmeans["Bernard"]) * cBernard,
         Nuria = (Nuria - cmeans["Nuria"]) * cNuria,
         Nerea = (Nerea - cmeans["Nerea"]) * cNerea,
         Carles = (Carles - cmeans["Carles"]) * cCarles,
         Ivan = (Ivan - cmeans["Ivan"]) * cIvan,
         Rachel = (Rachel - cmeans["Rachel"]) * cRachel,
         Nadia = (Nadia - cmeans["Nadia"]) * cNadia,
         Oriol = (Oriol - cmeans["Oriol"]) * cOriol,
         Valery = (Valery - cmeans["Valery"]) * cValery)

# row names was deleted to we need to create it again
corrdf$movie <- rownames(noVic)

# turn that move column back into row names
rownames(corrdf) <- corrdf[,21]
corrdf <- corrdf[,-21] # get rid of it now

# sum all the new rows
sumrows <- rowSums(corrdf, na.rm = T)

# find total pearson correlation
total_corr <- sum(abs(sumrows)) 
# 8.598467

# save average rating of Vic
VicA <- cmeans["Victoria"]

# the equation
final = VicA + (sumrows/total_corr)

# show top 5
Vic5 <- head(sort(final, decreasing = T), n=5)

# display
Vic5
```



# 4) Top similar critics
Return the top 5 critics that are most similar to Victoria. 

**Expected result**:
0.9449112, 'Rachel'
0.5976143, 'Ana' 
0.5789794, 'Oriol'
0.4925922, 'Maria'
0.4273247, 'Carles'

```{r fig.width=7, fig.height=6}
# use the correlation matrix created for question 1
critics_t_corr

# filter by Victoria column and remove Victoria row
Victoria <- critics_t_corr[-15, "Victoria"]

# sort by highest and limit to 5
head(sort(Victoria, decreasing = T), n=5)
```